---
title: "visualization EDA session 1"
author: "Caleigh Dwyer"
date: "2023-09-28"
output: github_document
---

#Viz 

```{r set up}
library(tidyverse)
library(ggridges)
```


```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
## using cached file: /Users/jeffgoldsmith/Library/Caches/org.R-project.R/R/rnoaa/noaa_ghcnd/USW00094728.dly
## date created (size, mb): 2023-09-19 15:41:55.07359 (8.524)
## file min/max dates: 1869-01-01 / 2023-09-30
## using cached file: /Users/jeffgoldsmith/Library/Caches/org.R-project.R/R/rnoaa/noaa_ghcnd/USW00022534.dly
## date created (size, mb): 2023-09-25 10:06:23.827176 (3.83)
## file min/max dates: 1949-10-01 / 2023-09-30
## using cached file: /Users/jeffgoldsmith/Library/Caches/org.R-project.R/R/rnoaa/noaa_ghcnd/USS0023B17S.dly
## date created (size, mb): 2023-09-19 15:42:03.139582 (0.994)
## file min/max dates: 1999-09-01 / 2023-09-30

weather_df
## # A tibble: 2,190 × 6
##    name           id          date        prcp  tmax  tmin
##    <chr>          <chr>       <date>     <dbl> <dbl> <dbl>
##  1 CentralPark_NY USW00094728 2021-01-01   157   4.4   0.6
##  2 CentralPark_NY USW00094728 2021-01-02    13  10.6   2.2
##  3 CentralPark_NY USW00094728 2021-01-03    56   3.3   1.1
##  4 CentralPark_NY USW00094728 2021-01-04     5   6.1   1.7
##  5 CentralPark_NY USW00094728 2021-01-05     0   5.6   2.2
##  6 CentralPark_NY USW00094728 2021-01-06     0   5     1.1
##  7 CentralPark_NY USW00094728 2021-01-07     0   5    -1  
##  8 CentralPark_NY USW00094728 2021-01-08     0   2.8  -2.7
##  9 CentralPark_NY USW00094728 2021-01-09     0   2.8  -4.3
## 10 CentralPark_NY USW00094728 2021-01-10     0   5    -1.6
## # ℹ 2,180 more rows
```


let's make a plot!

define aesthetic functions using aes. doesn't plot points without geom_point
```{r}

ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()

```

Pipes and stuff

use ggp_nyc_weather to save the plot as a named object

```{r}
weather_df |> 
  filter(name == "CentralPark_NY") |> 
  ggplot(aes(x=tmin, y = tmax)) + 
  geom_point ()

ggp_nyc_weather = 
  weather_df |> 
  filter(name == "CentralPark_NY") |> 
  ggplot(aes(x=tmin, y = tmax)) + 
  geom_point ()

```

## Fancy plot

you can also define aesthetics for point and smooth curves
You can set standard error (se) to false to take them out of the graph
set "alpha" aesthetics -- .3 means that it's 70% opaque
set "size" aesthetics -- .5 makes point biggers

```{r}
ggplot(weather_df, aes(x = tmin, y= tmax, color = name)) +
  geom_point(aes(color = name, alpha = .3, size = .5))+
  geom_smooth(se = FALSE)

```


##plot with facets

facet separates plot by variable

.~name says to organize plots by variable name

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  geom_smooth()+
  facet_grid(.~name)
```


let's try a different plot. temps are boring

```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name))+
  geom_point(aes(size=prcp),alpha = .3)+
  geom_smooth() +
  facet_grid(.~name)
```

try assigning a specific color....

```{r}
weather_df |> 
  filter(name == "CentralPark_NY") |> 
  ggplot(aes(x=date, y = tmax)) +
  geom_point(color = "blue")

##the below won't work! it thinks blue is a variable in the dataset
weather_df |> 
  filter(name == "CentralPark_NY") |> 
  ggplot(aes(x=date, y = tmax), color = "blue") +
  geom_point()

weather_df |> 
  filter(name != "CentralPark_NY") |> 
  ggplot(aes(x=date, y = tmax), color = name) +
  geom_point()

```

but for now, don't try to assign specific colors.


